---
title: "Homework 5"
author: "Tim Hauser"
output: github_document
---

## Initial setup

```{r}
library(tidyverse)
library(rvest)
library(p8105.datasets)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

The code chunk below imports the data in individual spreadsheets contained in `./data/zip_data/`. To do this, I create a dataframe that includes the list of all files in that directory and the complete path to each file. As a next step, I `map` over paths and import data using the `read_csv` function. Finally, I `unnest` the result of `map`.

```{r}
full_df = 
  tibble(
    files = list.files("data/zip_data/"),
    path = str_c("data/zip_data/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest()
```

The result of the previous code chunk isn't tidy -- data are wide rather than long, and some important variables are included as parts of others. The code chunk below tides the data using string manipulations on the file, converting from wide to long, and selecting relevant variables. 

```{r}
tidy_df = 
  full_df %>% 
  mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3)) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "outcome",
    names_prefix = "week_") %>% 
  mutate(week = as.numeric(week)) %>% 
  select(group, subj = files, week, outcome)
```

Finally, the code chunk below creates a plot showing individual data, faceted by group. 

```{r}
tidy_df %>% 
  ggplot(aes(x = week, y = outcome, group = subj, color = group)) + 
  geom_point() + 
  geom_path() + 
  facet_grid(~group)
```

This plot suggests high within-subject correlation -- subjects who start above average end up above average, and those that start below average end up below average. Subjects in the control group generally don't change over time, but those in the experiment group increase their outcome in a roughly linear way. 

## Problem 2

I downloaded the .csv file from github and added it to the data folder. The following code chunk imports the data into R as a dataframe:

```{r}
homicide_raw = read.csv("./data/homicide-data.csv") %>% 
  janitor::clean_names()
```

The homicides dataset was gathered and made public by The Washington Post on more than 52,000 criminal homicides over the past decade in 50 large U.S. cities. It contains ``r nrow(homicide_raw)`` observations of homicides and ``r ncol(homicide_raw)`` variables, namely ``r names(homicide_raw)``. The variables contain information about the date of the homicide, the victim (name, age, race, sex), the location of the homicide (city, state, lon, lat) and whether an arrest was made or not. 

Next, I created a city_state variable (e.g. “Baltimore, MD”) and status variable indicating whether a homicide was solved (those for which the disposition is “Closed by arrest”) or unsolved (those for which the disposition is “Closed without arrest” or “Open/No arrest”):

```{r}
homicide_df = homicide_raw %>% 
  mutate(city_state = str_c(city, ", ", state),
         status = case_when(
           disposition == "Closed without arrest" ~ "unsolved",
           disposition == "Open/No arrest" ~ "unsolved",
           disposition == "Closed by arrest" ~ "solved"
         )) %>% 
  relocate(city_state)
```

Next, I summarize within cities to obtain the total number of homicides and the number of unsolved homicides:

```{r}
homicide_df %>%
  group_by(city_state, status) %>%
  count(status) %>%
  spread(key = status, value = n) %>% 
  arrange(desc(unsolved)) %>% 
  knitr::kable()
```

The city with the most unsolved crimes (absolute, not as a %) is Chicago, IL.

Next, for the city of Baltimore, MD, I used the prop.test function to estimate the proportion of homicides that are unsolved; saved the output of prop.test as an R object, applied the broom::tidy to this object and pulled the estimated proportion and confidence intervals from the resulting tidy dataframe:

```{r}
baltimore_df = homicide_df %>%
  filter(city_state == "Baltimore, MD")

baltimore_unsolved = baltimore_df %>%   
  summarize(
    unsolved = sum(status == "unsolved"),
    n = n()
  )
  
baltimore_prop = 
  prop.test(
    x = baltimore_unsolved$unsolved,
    n = baltimore_unsolved$n) %>% 
  broom::tidy()
```

Baltimore has 64.5% unresolved homicides with a 95% CI of 62.8%-66.3%.


Next, I ran prop.test for each of the cities in the homicide dataset and extracted both the proportion of unsolved homicides and the confidence interval for each.

First, I created a function that calculates the proportion of unresolved crimes and a corresponding CI:

```{r}
prop_function = function(city_df) {
  
  city_unsolved =
    city_df %>%
    summarize(
      unsolved = sum(status == "unsolved"),
      n = n()
      )
  
  city_prop =
    prop.test(
      x = city_unsolved$unsolved,
      n = city_unsolved$n)
  
  return(city_prop)
}
```

Second, I created a nested dataframe:

```{r}
nested_df = homicide_df %>%
  nest(data = uid:status) %>% 
  mutate(
    test_results = map(data, prop_function),
    tidy_results = map(test_results, broom::tidy)
  ) %>% 
  select(city_state, tidy_results) %>% 
  unnest(tidy_results) %>% 
  select(city_state, estimate, starts_with("conf"))
```

Lastly, I created a plot that shows the estimates and CIs for each city using geom_errorbar to add error bars based on the upper and lower limits. I organized cities according to the proportion of unsolved homicides:

```{r}
nested_df %>%
  mutate(city_state = fct_reorder(city_state, estimate)) %>%
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Proportion of unresolved homicides including 95% CI by city",
       x = "City, State",
       y = "% unresolved homicides") 
```


## Problem 3

First, I am setting key design elements and variables to start the problem

```{r}
n = 30
sd = 5
true_mean = 0:6
dataset = function(x) {
  rnorm(n, x, sd)}
output = vector("list", length = 5000)
inter = data.frame(matrix(ncol = 3, nrow = 5000)) 
colnames(inter) = c('true_mean', 'sample_mean', 'p-value')
```

Creating initial function and final list with all 5000 datasets for each true_mean

```{r}
combine_function = function(y){
  inter2 = list()
  for (i in 1:5000) {
    output[[i]] = dataset(y)
    to_be_deleted = broom::tidy(t.test(output[[i]]))
    inter[i,1] = y
    inter[i,2] = to_be_deleted[[1]]
    inter[i,3] = to_be_deleted[[3]]}
  inter2 = rbind(inter2, inter)}

final_list = map(true_mean, combine_function)
```

### Converting list data into a single dataframe that can be utilized for subsequent visualization
```{r}
df_final = data.frame(c(1:5000),final_list[[1]])
for (i in 2:7) {
  df = data.frame(c(1:5000),final_list[[i]])
  df_final = rbind(df_final, df)
  }
df_final = df_final %>%
  mutate(reject = ifelse(p.value < 0.05, TRUE, FALSE))
```

### Graph #1 - Showing the proportion of times the null was rejected

```{r}
reject_plot = df_final %>%
  group_by(true_mean) %>% 
  summarize(power = sum(reject)/5000) %>% 
  ggplot(aes(x = true_mean, y = power)) +
  geom_point(aes(colour = true_mean)) +
  geom_line(alpha = 0.3) +
  labs(x = "True Mean", 
       y = "Proportion of times  null was rejected") +
  scale_x_continuous(breaks = seq(0,6,by = 1)) +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,by = 0.2))
reject_plot
```

The graph clearly shows that as effect size increases, the power also increases. This association is in accordance with basic theory.


### Graph #2 - Comparing estimated and true Means

The following two graphs compare estimated and true means. 

-   The left graph shows the average estimate of the mean on the y axis and the true mean on the x axis for all observations.
-   The right graph shows the average estimate of the mean on the y axis and the true mean on the x axis for samples for which the null was rejected.

```{r}
Comparison_plot = df_final %>%
  group_by(true_mean) %>%
  summarize(mean_sample = mean(sample_mean)) %>%
  ggplot(aes(x = true_mean, y = mean_sample)) +
  geom_point(aes(colour = true_mean)) +
  labs(title = "Total Sample",
       x = "True Mean", 
       y = "Sample Mean") +
  scale_x_continuous(limits = c(0,6), breaks = seq(0,6,1)) +
  scale_y_continuous(limits = c(-2,7), breaks = seq(-2,7,1)) +
  theme(plot.title = element_text(hjust = 0.5))

Comparison_reject = df_final %>%
  group_by(true_mean) %>%
  filter(p.value < 0.05) %>%
  summarize(mean_sample = mean(sample_mean)) %>%
  ggplot(aes(x = true_mean, y = mean_sample)) +
  geom_point(aes(colour = true_mean)) +
  labs(title = "Rejected Null Sample",
       x = "True Mean", 
       y = "Sample Mean") +
  scale_x_continuous(limits = c(0,6), breaks = seq(0,6,1)) +
  scale_y_continuous(limits = c(-2,7), breaks = seq(-2,7,1)) +
  theme(plot.title = element_text(hjust = 0.5))

Comparison_plot + Comparison_reject
```

As the effect size increases, the sample average of the mean starts to be closer to the true mean. This makes sense, as a larger effect size increases the power.