---
title: "Homework 6"
author: "Tim Hauser"
output: github_document
---

## Initial setup

```{r}
library(tidyverse)
library(viridis)
library(modelr)


#library(p8105.datasets)
#library(moderndive)

set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Loading the data:

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

In the following problem, we’ll focus on a simple linear regression with tmax as the response and tmin as the predictor, and are interested in the distribution of two quantities estimated from these data: r^2 and log(β^0∗β^1. We use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. 

First, for r^2: To obtain a distribution for r^2, we'll follow basically the same procedure we used for regression coefficients: draw bootstrap samples; the a model to each; extract the value I'm concerned with; and summarize. Here, we'll use `modelr::bootstrap` to draw the samples and `broom::glance` to produce `r.squared` values. 

```{r}
weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  summarize(
    mean_r_sq = mean(r.squared),
    sd_r_sq = sd(r.squared),
    ci_lower = quantile(r.squared, 0.025), 
    ci_upper = quantile(r.squared, 0.975)
    )
```

Then, plotting the distrbution of r^2:

```{r}
weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density()
```

In this example, the r^2 value is high, and the upper bound at 1 may be a cause for the generally skewed shape of the distribution. If we wanted to construct a confidence interval for r^2, we could take the 2.5% and 97.5% quantiles of the estimates across bootstrap samples. However, because the shape isn't symmetric, using the mean +/- 1.96 times the standard error probably wouldn't work well.


Next, we move on to computing log(β^0∗β^1: We can produce a distribution for log(β^0∗β^1 using a similar approach, with a bit more wrangling before we make our plot.

```{r}
weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  summarize(
    mean_log_b0b1 = mean(log_b0b1),
    sd_log_b0b1 = sd(log_b0b1),
    ci_lower = quantile(log_b0b1, 0.025), 
    ci_upper = quantile(log_b0b1, 0.975)
  )
```

Now we plot the distribution of log(β^0∗β^1:

```{r}
weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  ggplot(aes(x = log_b0b1)) + geom_density()
```

As with r^2, this distribution is somewhat skewed and has some outliers. 

## Problem 2









